---
title: "hg_ah"
output: html_notebook
---

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(scales)
library(units)
```

```{r}
property_details <- read.csv(file = "../data/property_details.csv")
LIHTC <- read.csv(file = "../data/LIHTC_updated.csv")
filtered_sales <- read.csv(file = "../data/filtered_sales.csv")
barnes <- read.csv(file = "../data/barnes.csv")
asssesments <- read.csv(file = "../data/assessments.csv")
```

Filter barnes and LIHTC to cordnates and year put into services, so we can 
stack them.
```{r}
LIHTC <- LIHTC %>% 
  select(YR_PIS, LATITUDE, LONGITUDE, HUD_ID) %>% 
  rename(id = HUD_ID, year = YR_PIS, lat = LATITUDE, lng = LONGITUDE) %>% 
  mutate(type = 'LIHTC')

barnes <- barnes %>% 
  select(Barnes.Year, lat, lng) %>% 
  mutate(bid = paste0('B', row_number())) %>% 
  rename(id = bid, year = Barnes.Year) %>% 
  mutate(type = 'barnes')

all_ah_info <- LIHTC %>% 
  bind_rows(barnes) %>% 
  filter(year >= 2000 & year < 5000)
```

convert the sale date chr to sales year int
```{r}
filtered_sales$ownerdate <- str_sub(filtered_sales$ownerdate, 1, 4)
```

```{r}
filtered_sales <- transform(filtered_sales, ownerdate = as.numeric(ownerdate))
```

Split the centroid column into lat and lng.
```{r}
property_details <- property_details %>% 
  separate(centroid, c('lng', 'lat'), ',')

property_details$lng <- str_remove(property_details$lng, '[(]')
property_details$lat <- str_remove(property_details$lat, '[)]')
```

merge the property details with sales data
```{r}
property_details <- inner_join(filtered_sales, property_details, by = 'apn')
```

converting cords to a geometry
```{r}
property_details <- property_details %>% 
  st_as_sf(coords = c("lat", "lng"), crs = 4326, agr = "identity") %>% 
  rename(home_loc = geometry)

all_ah_info <- all_ah_info %>% 
  st_as_sf(coords = c("lat", "lng"), crs = 4326, agr = "identity")
```

find the nearest ah to each home.
```{r}
nearest_ah <- bind_cols(property_details, 
          all_ah_info[st_nearest_feature(property_details, all_ah_info), ])
```

```{r}
nearest_ah <- nearest_ah %>% 
  mutate(dist = st_distance(nearest_ah$home_loc, nearest_ah$geometry, by_element = T))
```

```{r}
nearest_ah$dist <- drop_units(nearest_ah$dist/1609.34) #convert to Mi
```

filtering down to 1 mi
```{r}
nearest_ah <- nearest_ah %>% 
  filter(dist < 1)
```

For each remaining home, calculate a new column called "group", which is defined according to the following rules. Hint: Use the case_when function to do this.
  "pre" - for homes where the distance is less than half a mile and whose sale date was 2-5 years prior to the input year
  "mid" - for homes where the distance is less than half a mile and whose sale date was 0-2 years prior to the input year
  "post" - for homes where the distance is less than half a mile and whose sale date was after the input year
  "outside" - for homes where the distance is more than half a mile and whose sale date was no more than 5 years prior to the input year
  "other" - All other rows
```{r}
nearest_ah <- nearest_ah %>% 
  mutate(
    group = case_when(
    dist <= .5 & year - ownerdate >= 2 & year - ownerdate <= 5 ~ 'pre',
    dist <= .5 & year - ownerdate < 2 & year - ownerdate == 0 ~ 'mid',
    dist <= .5 & ownerdate - year > 0 ~ 'post',
    dist > .5 & year - ownerdate <= 5 ~ 'outside',
    T ~ 'other'
    )
  ) %>% 
  filter(group != 'other')
```

creating a tpost that is years sale occured after the houseing development
```{r}
nearest_ah <- nearest_ah %>% 
  mutate(
    tpost = if_else(group == 'post', ownerdate - year, 0)
    )
```

create an age column
```{r}
nearest_ah <- nearest_ah %>% 
  mutate(age = ownerdate - year_built) %>% 
  filter(age >= 0) %>% 
  drop_na(age)
```

filter down to 5 years before or after
```{r}
nearest_ah <- nearest_ah %>% 
  filter(year - ownerdate <= 5 & ownerdate - year <= 5) %>% 
  distinct() 
```

Forming a liner model for the data
```{r}
simple_regression_model <- lm(amount ~ square_footage + age + group + ownerdate + tract, data = nearest_ah)

summary(simple_regression_model)
```

```{r}
nearest_ah %>%
  mutate(predicted = predict(simple_regression_model, newdata = nearest_ah)) %>%
  ggplot(aes(x = amount, y = predicted)) +
    geom_point(na.rm = TRUE)
```

redoing the linear model but this time with the log of amount
```{r}
nearest_ah <- nearest_ah %>% 
  mutate(group = factor(group), tract = factor(tract))


simple_regression_model_log <- lm(log(amount) ~ square_footage + age + relevel(group, ref = 'outside') + ownerdate + tpost + tract, data = nearest_ah)

summary(simple_regression_model_log)
```

```{r}
nearest_ah %>%
  mutate(predicted = predict(simple_regression_model_log, newdata = nearest_ah)) %>%
  ggplot(aes(x = log(amount), y = predicted)) +
    geom_point(na.rm = TRUE)
```

```{r}
anova(simple_regression_model_log)
```


```{r}
nearest_ah$land_area <- str_sub(nearest_ah$land_area, 1, 4)
```

```{r}
nearest_ah <- transform(nearest_ah, land_area = as.numeric(land_area))
```

```{r}
simple_regression_model_log_la <- lm(log(amount) ~ square_footage + age + group + ownerdate + tract + land_area, data = nearest_ah)

summary(simple_regression_model_log_la)
```

